### 2021-11-10 NSF2021/Description/specificaim2.tex
working on double well results

\label{subsec:specific_aim_2}
\begin{wrapfigure}[14]{r}{0.61\textwidth}
  \centering{\includegraphics[width=0.60\textwidth]{Figures/domain.pdf}}
    \vspace{-15pt}
    \caption{\label{fig:flow_map} \footnotesize A collection of 3D
    amphiphilic particles suspended in a solvent. The solvent velocity
    satisfies the Stokes equations, and the hydrophobic forces depend on
    the solvent activity that satisfies the screened Laplace equation.
    The geometry is updated by solving the mobility problem.} 
\end{wrapfigure}
The HAP model requires solving exterior Dirichlet problems for the order
parameter~\eqref{eq:SL} and the Stokes
equations~\eqref{eq:stokes1}--\eqref{eq:stokes2}. If discretizing these
equations with stencil-based numerical methods, the computational domain
must be truncated, the fluid volume must be discretized, and artificial
boundary conditions must be imposed. The artificial boundary conditions
introduce additional error, and discretizing the volume results in an
excessively large system of equations. Moreover, it is difficult to
obtain high-order discretizations when the boundary surfaces are
irregular.

@@

Significant challenges to solving BIEs include solving largely stem from
solving dense linear systems. \S\ref{subsec:NumericalIssues}
describes how we will combine preconditioning strategies with fast
summation methods to accelerate the time to solution,
\S\ref{subsec:timeStepping} describes adaptive time stepping methods,
and \S\ref{subsec:AC} describes how the non-linear Allen-Cahn
equation~\eqref{eq:SL} will be iteratively solved using BIE
formulations.

\subsubsection{Iterative methods for the Allen-Cahn HAP model}
\label{subsec:AC}
In the PIs collaboration, they have developed numerical methods to
solve~\eqref{eq:SL} when $f(u) = u^2$. With this choice of $f(u)$, the
resulting PDE is the linear screened Laplace equation which is
efficiently solved with a double-layer potential representation. For a
more general $f$, such as a double well potential, we will interpret the
solution of~\eqref{eq:SL} as a root finding problem for the operator
\begin{align*}
  F[u] = -\rho^2 \Delta u + f'(u).
\end{align*}
This non-linear PDE will be solved with Newton's method
\begin{align*}
  u^{N+1} = u^{N} - dF^{-1} F(u^N),
\end{align*}
where $dF = -\rho^2 \Delta + f''(u^N)$ is the Fr\'{e}chet derivative of
$F$. As expected, the biggest challenge amounts to solving
\begin{align}
  \label{eq:NewtonUpdate}
  - \rho^2 \Delta v + f''(u^{N}) v = F[u^N].
\end{align}
For a general double well potential $f(u)$, this is a variable
coefficient elliptic PDE, and developing an integral equation
formulation is challenging. We propose to instead use an inexact
Newton's method where $dF^{-1}$ is defined in terms of the piecewise
quadratic approximation
\begin{align*}
  f(u) \approx bu + \begin{cases}
    a(u - U_1)^2, &\mbox{if } u \leq (U_0 + U_1)/2 \\
    (u - U_2)^2, &\mbox{if } u > (U_0 + U_1)/2.
  \end{cases}
\end{align*}
In this case, the update step of Newton's method~\eqref{eq:NewtonUpdate}
amounts to solving the piecewise constant PDE
\begin{align}
  \label{eq:screenedPoisson}
  -\rho^2 \Delta v + 
  \left(
  \begin{cases}
    2a, &\mbox{if } u^{N} \leq (U_0 + U_1)/2 \\
    2, &\mbox{if } u^{N} > (U_0 + U_1)/2
  \end{cases}
  \right)
  v = F[u^N],
\end{align}
which is a screened Poisson equation with piecewise a constant screening
coefficient. Solutions of the screened Poisson equation can be written
in terms of volume and layer potentials. We first define the level set
$\Gamma = \{\xx \:|\: \uu(\xx) = (U_0 + U_1)/2 \}$ which separates the
domain $\Omega$ into two regions. Then, assuming that the particles are
all contained in the hydrophobic region, a solution of the screened
Laplace equation (i.e.~equation~\eqref{eq:screenedPoisson} when $F[u^N]
= 0$) is
\begin{align*}
  u(\xx) = \frac{1}{2\pi} \int_{\bd\Omega} \pderiv{}{\nn_\yy} K_0 \left(
    \frac{|\xx - \yy|}{\rho}\right) \sigma(\yy) ds_\yy + 
    \frac{1}{2\pi} \int_{\Gamma} K_0 \left(
    \frac{|\xx - \yy|}{\rho} \right) \eta(\yy) ds_\yy,
\end{align*}

>>>>

\label{subsec:specific_aim_2}
\begin{wrapfigure}[14]{r}{0.61\textwidth}
  \centering{\includegraphics[width=0.60\textwidth]{Figures/domain.pdf}}
    \vspace{-15pt}
    \caption{\label{fig:flow_map} \footnotesize A collection of 3D
    amphiphilic particles suspended in a solvent. The solvent velocity
    satisfies the Stokes equations, and the hydrophobic forces depend on
    the solvent activity that satisfies the screened Laplace equation.
    The geometry is updated by solving the mobility problem.} 
\end{wrapfigure}
The HAP model requires solving exterior Dirichlet problems for the order
parameter~\eqref{eq:SL} and the Stokes
equations~\eqref{eq:stokes1}--\eqref{eq:stokes3}. If discretizing these
equations with stencil-based numerical methods, the computational domain
must be truncated, the fluid volume must be discretized, and artificial
boundary conditions must be imposed. The artificial boundary conditions
introduce additional error, and discretizing the volume results in an
excessively large system of equations. Moreover, it is difficult to
obtain high-order discretizations when the boundary surfaces are
irregular.

@@

Significant challenges to solving BIEs include solving largely stem from
solving dense linear systems. First, \S\ref{subsec:AC} describes how the
non-linear Allen-Cahn equation~\eqref{eq:SL} will be iteratively solved
using BIE formulations. Then, \S\ref{subsec:NumericalIssues} describes
how we will combine preconditioning strategies with fast summation
methods to accelerate the time to solution, and
\S\ref{subsec:timeStepping} describes adaptive time stepping methods,

\subsubsection{Iterative methods for the Allen-Cahn HAP model}
\label{subsec:AC}

\begin{wrapfigure}[]{l}{1.8in}
  \centering
  \input{figures/doublewell.tikz}
  \caption{\label{fig:well} $U_1 = 0.5$, $U_2 = 2$, $a = 1$, $b = 0.4$.}
\end{wrapfigure}
In the PIs collaboration, they have developed numerical methods to
solve~\eqref{eq:SL} when $f(u) = u^2$. With this choice of $f(u)$, the
resulting PDE is the linear screened Laplace equation which is
efficiently solved with a double-layer potential representation. For a
more general $f$, such as a double well potential, we will interpret the
solution of~\eqref{eq:SL} as a root finding problem for the operator
\begin{align}
  \label{eq:F}
  F[u] = -\rho^2 \Delta u + f'(u).
\end{align}
This non-linear PDE will be solved with Newton's method
\begin{align*}
  u^{N+1} = u^{N} - \alpha dF^{-1} F(u^N),
\end{align*}
where $dF = -\rho^2 \Delta + f''(u^N)$ is the Fr\'{e}chet derivative of
$F$, and $\alpha$ is the step size which will be determined using
backtracking. As expected, the biggest challenge amounts to solving
\begin{align}
  \label{eq:NewtonUpdate}
  - \rho^2 \Delta v + f''(u^{N}) v = F[u^N].
\end{align}
For a general double well potential $f(u)$, this is a variable
coefficient elliptic PDE, and developing an integral equation
formulation is challenging. We propose to instead use an inexact
Newton's method where $dF^{-1}$ is defined in terms of the piecewise
quadratic approximation
\begin{align*}
  f(u) \approx \tilde{f}(u) = bu + \begin{cases}
    a(u - U_1)^2, &\mbox{if } u \leq (U_0 + U_1)/2 \\
    (u - U_2)^2, &\mbox{if } u > (U_0 + U_1)/2.
  \end{cases}
\end{align*}
In this case, the update step of Newton's method~\eqref{eq:NewtonUpdate}
amounts to solving the PDE
\begin{align}
  \label{eq:screenedPoisson}
  -\rho^2 \Delta v + \tilde{f}''(u^{N})v = F[u^N],
\end{align}
\begin{wrapfigure}[]{r}{1.8in}
  \centering
  \input{figures/residuals.tikz}
  \caption{\label{fig:residuals} $U_1 = 0.5$, $U_2 = 2$, $a = 1$, $b =
  0.4$.}
\end{wrapfigure}
which is a screened Poisson equation with piecewise a constant screening
coefficient. We test the efficacy of Newton's method applied
to~\eqref{eq:F} by considering its one-dimensional version in $(0,1)$
with the Dirichlet boundary conditions $u(0) = U_0$ and $u(1) = U_1$.
These parameters result in the double well potential illustrated in
Figure~\ref{fig:well}(a), and the inexact Newton's method converges to
the solution in Figure~\ref{fig:well}(b). Figure~\ref{fig:residuals}
shows the max norm of the residual at each Newton iterate when using
the exact second derivative (red curve) and the piecewise constant
approximate (blue curve).

Solutions of the screened Poisson equation can be written
in terms of volume and layer potentials. We first define the level set
$\Gamma = \{\xx \:|\: \uu(\xx) = (U_0 + U_1)/2 \}$ which separates the
domain $\Omega$ into two regions. Then, assuming that the particles are
all contained in the hydrophobic region, a solution of the screened
Laplace equation (i.e.~equation~\eqref{eq:screenedPoisson} when $F[u^N]
= 0$) is
\begin{align*}
  u(\xx) = \frac{1}{2\pi} \int_{\bd\Omega} \pderiv{}{\nn_\yy} K_0 \left(
    \frac{|\xx - \yy|}{\rho}\right) \sigma(\yy) ds_\yy + 
    \frac{1}{2\pi} \int_{\Gamma} K_0 \left(
    \frac{|\xx - \yy|}{\rho} \right) \eta(\yy) ds_\yy,
\end{align*}
